Implications of animal movement rate variation in spatial patterns of seed dispersal

### 2021-11-23 Ch1_movement_rates/Ch1_Figures.Rmd
I'm going to say, figure 4 is ready.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(fitdistrplus)
library(Hmisc)
library(gtsummary)
```

@@

  seed.data %>% 
    drop_na(s.id) %>% 
    mutate(dispcat = ifelse(disp <150, "non", 
                            ifelse(disp <500, "med", "ldd"))) %>% 
    #slice_head(., n = s.size) %>% 
    sample_n(s.size) %>% 
      ggplot(., aes(x = xloc, y = yloc, color = dispcat)) +
      geom_point(alpha = lalpha) +
      scale_color_manual(values = c(my.cols1[6], my.cols1[7], "black")) +
      geom_text(label = scenario, x = -1100, y = -1100, color = "black") +
      coord_equal(xlim = c(-xylims, xylims), ylim = c(-xylims, xylims)) +
      theme_bw() +
      theme(legend.position = "none",
            axis.title = element_blank()) +
       NULL -> seedlocplot

@@

          summ.data %>% 
    #drop_na(s.id) %>% 
    mutate(dispcat = ifelse(av.disp <500, "non", "ldd")) %>% 
    #slice_head(., n = s.size) %>% 
    sample_n(s.size*5) %>% 
      ggplot(., aes(x = x, y = y, color = dispcat)) +
      geom_point(alpha = lalpha) +
      labs(subtitle = scenario) +
      scale_color_manual(values = c(my.cols1[6], my.cols1[7])) +
      coord_equal(xlim = c(-xylims, xylims), ylim = c(-xylims, xylims)) +
      theme_bw() +
      theme(legend.position = "none",
            axis.title = element_blank()) +
       NULL -> summaryplot

@@

seedplots2d <- plot_grid(A, B, C, ncol = 3, labels = "A")
#seedplots2d

@@

```{r fig.height=7.5, fig.width=9}
plot_grid(seedplots2d, disp2_plot, nrow = 2)

ggsave2(filename = "Ch1_movement_rates/Figures/Figure4.png")

@@

### EVD generalized pareto results

@@

```{r}
library(extRemes)

@@

```{r fig.show='hide'}

nint <- 70 # number of intervals. how many thresholds to try
r <- c(0, 700) # range over which to try the thresholds
nsamps <- 500 # random sample form the distances

cp_threshplot <- threshrange.plot(cp.evd$disp, r = r, type = "GP", nint = nint)

@@

```{r}
as.data.frame(cp_threshplot) %>%
  mutate(u.i = seq(r[1], r[2], length.out = nint)) -> cp_threshplot

@@

```{r}
a <- read.table("Ch1_movement_rates/dispdistancestest.txt")
a

threshrange.plot(a$V1, type = "GP", nint = 100, r = c(0,700))
c <- mrlplot(a$V1)
mrlplot(a$V1)

max(c[,3])

as.data.frame(c) -> d

model <- lm(d$`Mean Excess`~1)
confint(model, level = 0.9)

b <- decluster(a$V1, threshold = 170)

threshrange.plot(b)

fit <- fevd(b, threshold = 170, type = "GP")
plot(fit)
fit$results

1-pevd(1000,threshold = 170, scale= fit$results$par[1], shape = fit$results$par[2], type = "GP")

# OR

# To find the probability of exceeding 1000, 5000, or 10,000 meters based on the fitted ged 
dist_range <- c(1000, 5000, 10000)
pextRemes(fit, dist_range, lower.tail = FALSE)

>>>>

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(fitdistrplus)
library(Hmisc)
library(gtsummary)
library(extRemes)
```

@@

  # seed.data %>% 
  #   drop_na(s.id) %>% 
  #   mutate(dispcat = ifelse(disp <150, "non", 
  #                           ifelse(disp <500, "med", "ldd"))) %>% 
  #   #slice_head(., n = s.size) %>% 
  #   sample_n(s.size) %>% 
  #     ggplot(., aes(x = xloc, y = yloc, color = dispcat)) +
  #     geom_point(alpha = lalpha) +
  #     scale_color_manual(values = c(my.cols1[6], my.cols1[7], "black")) +
  #     geom_text(label = scenario, x = -1100, y = -1100, color = "black") +
  #     coord_equal(xlim = c(-xylims, xylims), ylim = c(-xylims, xylims)) +
  #     theme_bw() +
  #     theme(legend.position = "none",
  #           axis.title = element_blank()) +
  #      NULL -> seedlocplot

  seed.data %>% 
    drop_na(s.id) %>% 
    mutate(LDD = ifelse(disp <500, NA, "Over 500")) %>% 
    sample_n(s.size) %>% 
      ggplot(., aes(x = xloc, 
                    y = yloc
                    , color = LDD
                    )) +
      geom_hex()+
      scale_fill_gradient(low = "grey", high = "black") +
      #geom_point(alpha = lalpha) +
      scale_color_manual(values = c(my.cols1[6], my.cols1[7])) +
      geom_text(label = scenario, x = -1100, y = -1100, color = "black") +
      coord_equal(xlim = c(-xylims, xylims), ylim = c(-xylims, xylims)) +
      theme_bw() +
      theme(legend.position = "none",
            axis.title = element_blank()) +
       NULL -> seedlocplot

@@

          summ.data %>% 
    #drop_na(s.id) %>% 
    mutate(dispcat = ifelse(av.disp <500, "non", "Over 500")) %>% 
    #slice_head(., n = s.size) %>% 
    sample_n(s.size*5) %>% 
      ggplot(., aes(x = x, y = y, color = dispcat)) +
      geom_point(alpha = lalpha) +
      labs(subtitle = scenario) +
      scale_color_manual(values = c(my.cols1[6], my.cols1[7])) +
      coord_equal(xlim = c(-xylims, xylims), ylim = c(-xylims, xylims)) +
      theme_bw() +
      theme(legend.position = "none",
            axis.title = element_blank()) +
       NULL -> summaryplot

@@

prow <- plot_grid(A, B, C, ncol = 3, labels = "A")

legend <- get_legend(C + scale_fill_gradient(low = "grey", high = "black",
                                             name = "Count",
                          guide = guide_colorbar(barwidth = 15, 
                                                barheight = 0.5,
                                                title.position = "left",
                                                title.vjust = 1.1,
                                                 #label.position = "left",
                                                 nbin = 100)) +
  scale_color_manual(values = c(LDD = my.cols1[6]), name = "")+
    guides(color = guide_legend(override.aes = list(fill = my.cols1[6]))) +
    theme(legend.position = "bottom"))

seedplots2d <- plot_grid(prow, legend, rel_heights = c(1,0.1), ncol = 1)
#seedplots2d

@@

```{r fig.height=7.5, fig.width=9}
plot_grid(seedplots2d, NULL, disp2_plot, nrow = 3, rel_heights = c(1,0.2,1))

#ggsave2(filename = "Ch1_movement_rates/Figures/Figure4.png")

@@

### FIGURE 6 EVD generalized pareto results

@@

```{r results='hide'}

# subset data ---------------------------------

nsamps <- 10000
itera <- 1000

pp.df %>%
  drop_na(., s.id) %>%
  sample_n(., nsamps) -> pp.evd

np.df %>%
  drop_na(., s.id) %>%
  sample_n(., nsamps) -> np.evd

cp.evd %>%
  mutate(model = toupper(model)) %>%
  bind_rows(pp.evd %>% mutate(model = toupper(model))) %>%
  bind_rows(np.evd %>% mutate(model = toupper(model))) %>%
  group_by(model) %>%
  summarise(q75 = quantile(disp, 0.75),
            q90 = quantile(disp, 0.9),
            q99 = quantile(disp, 0.99),
            max = max(disp))

# Find threshold -------------------------------------
## CP ----------------------------

r <- c(0, 600)

nint <- 100
# r <- quantile(cp.evd$disp, c(0.75, 0.99))
u.i <- matrix(seq(r[1],r[2],,nint), ncol=1)

cp.mrl <- mrlplot(cp.evd$disp)
lm.cp <- lm(cp.mrl[,2]~1)
tval.cp <- confint(lm.cp, level = 0.9)[2]

thresh.cp <- threshrange.plot(cp.evd$disp, r = r, type = "GP", nint = nint) %>%
  as.data.frame() %>%
  mutate(u.i = u.i)

## PP ----------------------------

nint <- 100
# r <- quantile(pp.evd$disp, c(0.75, 0.99))
u.i <- matrix(seq(r[1],r[2],,nint), ncol=1)

pp.mrl <- mrlplot(pp.evd$disp)
lm.pp <- lm(pp.mrl[,2]~1)
tval.pp <- confint(lm.pp, level = 0.9)[2]

thresh.pp <- threshrange.plot(pp.evd$disp, r = r, type = "GP", nint = nint) %>%
  as.data.frame() %>%
  mutate(u.i = u.i)

## NP ----------------------------

nint <- 100
# r <- quantile(pp.evd$disp, c(0.75, 0.99))
u.i <- matrix(seq(r[1],r[2],,nint), ncol=1)

np.mrl <- mrlplot(np.evd$disp)
lm.np <- lm(np.mrl[,2]~1)
tval.np <- confint(lm.np, level = 0.9)[2]

thresh.np <- threshrange.plot(np.evd$disp, r = r, type = "GP", nint = nint) %>%
  as.data.frame() %>%
  mutate(u.i = u.i)

@@

```{r fig.width=12, fig.height=3}
# Threshold figures

threshplot_fx <- function(thresh_data, th, r, model){

  th <- round(th)

  thresh_data %>%
    ggplot(., aes(y = t.scale, x = u.i)) +
    geom_point(shape = 1, size = 2) +
    # geom_line(linetype = "dashed") +
    geom_linerange(aes(x = u.i, ymin = low.t.scale, ymax = up.t.scale)) +
    labs(title = paste(model, "threshold =", th, "m"), 
         x = "Threshold in meters", y = "Scale") +
    theme_bw() +
    geom_vline(xintercept = th, color = "red", linetype = "dashed") +
    scale_x_continuous(limits = r, n.breaks = 7 ) +
    # scale_x_continuous(name = NULL, labels = NULL, breaks = NULL) +
    # theme(axis.title.x = element_blank(),
    #       axis.ticks.x = element_blank(),
    #       axis.text.x = element_blank()) +
    # coord_cartesian(xlim = c(200, 550)) +
    NULL -> thresh_scale

  thresh_data %>%
    ggplot(., aes(y = shape, x = u.i)) +
    geom_point(shape = 1, size = 2) +
    geom_linerange(aes(x = u.i, ymin = low.shape, ymax = up.shape)) +
    labs(x = "Threshold in meters", y = "Shape") +
    theme_bw() +
    geom_vline(xintercept = th, color = "red", linetype = "dashed") +
    scale_x_continuous(limits = r, n.breaks = 7 ) +
    # coord_cartesian(xlim = c(200, 550)) +
    NULL -> thresh_shape

  plot_grid(thresh_scale, thresh_shape, nrow = 2)

}

plot.cp <- threshplot_fx(thresh_data = thresh.cp, th = tval.cp, r = r, model = "CP")
plot.pp <- threshplot_fx(thresh_data = thresh.pp, th = tval.pp, r = r, model = "PP")
plot.np <- threshplot_fx(thresh_data = thresh.np, th = tval.np, r = r, model = "NP")

plot_grid(plot.cp, plot.pp, plot.np, ncol = 3)
ggsave2(filename = "Ch1_movement_rates/Figures/Figure6.png")

@@

## Fit values

evd.fit.cp <-fevd(cp.evd$disp, threshold = tval.cp, type = "GP")
evd.fit.pp <- fevd(pp.evd$disp, threshold = tval.cp, type = "GP")
evd.fit.np <- fevd(np.evd$disp, threshold = tval.cp, type = "GP")

summary(evd.fit.cp)
summary(evd.fit.pp)
summary(evd.fit.np)

@@

Find the conditional probabilities. Probability of exceeding a distance x, based on the fitted distribution

# To find the probability of exceeding 1000, 5000, or 10,000 meters based on the fitted ged 
dist_range <- c(200, 500, 1000, 5000, 10000)

pextRemes(evd.fit.cp, dist_range, lower.tail = FALSE)
pextRemes(evd.fit.pp, dist_range, lower.tail = FALSE)
pextRemes(evd.fit.np, dist_range, lower.tail = FALSE)
